---
title: "Lena-pred Data Visualization"
output: html_document
---

## Set up

```{r libraries and config, warning=FALSE, include=FALSE}
source(here::here("code/00_config/lena-pred-libraries.R"))
source(here("code/00_config/lena-pred-config.R"))

# lstm  results
preds_files <- list.files(config_obj$paths_config$lstm_preds_path)

d <- preds_files %>% 
  map(read_lena_pred_data,
      config_obj, 
      f_type = "rds", 
      is_pitch = FALSE) %>% 
  setNames(config_obj$exp_config$runs) 

d <- d %>% future_map2_dfr(.f = process_exp_run, .y = names(.))
d$speech_register <- factor(d$speech_register) %>% fct_rev()

# persistence model results
persist_file <- list.files(config_obj$paths_config$lstm_sum_path, 
                           pattern = '*persist=*')

d_persist <- read_lena_pred_data(persist_file, config_obj, f_type = "csv")

if (config_obj$exp_config$dataset_name == "IDSLabel") {
  d <- d %>% mutate(speaker_id = str_split(seg_id, "_", simplify = TRUE)[1])  
}
```

## Baseline persistence model results. 

```{r}
ms_persist <- d_persist %>% 
  filter(!(is.na(speaker_id)), 
         prop_cds == 0.5) %>% 
  group_by(speech_register, speaker_id, seg_id) %>% 
  summarise(m = mean(m, na.rm = T)) %>% 
  group_by(speaker_id, speech_register) %>% 
  summarise(m = mean(m, na.rm = T)) %>% 
  group_by(speech_register) %>% 
  tidyboot_mean(col = m, na.rm = T) %>% 
  mutate(model = "persistence") %>% 
  rename(m = mean) %>% 
  select(-n, -empirical_stat)
```

## ADS vs. CDS average predictability

Replicate the key analysis in Rasanen et al. (2018), which compared the average predictability of ADS vs. CDS.

First we aggregate the results for each run of the experiment for only the 50/50 split of CDS in the training data.

```{r}
d_analysis <- d %>% 
  filter(prop_cds_train == 0.5, 
         is_target_cluster == TRUE) %>% 
  mutate(exp_run_id = str_remove(exp_run_id, "run"),
         duration_ms = as.numeric(duration_ms),
         duration_sec = duration_ms / 1000,
         time_bin_id_num = as.numeric(time_bin_id)) 

 d_nest <- d_analysis %>% 
   group_by(fold_id, exp_run_id) %>% 
   nest() %>% 
   mutate(results = map(data, analyze_one_run, 
                        metric = m_acc))

 # aggregate across runs
 ms_lstm <- d_nest %>% 
   unnest(results) %>%
   group_by(speech_register, exp_run_id) %>% 
   summarise(m = mean(empirical_stat), 
             ci_lower = mean(ci_lower),
             ci_upper = mean(ci_upper)) %>% 
   group_by(speech_register) %>% 
   summarise(m = mean(m), 
             ci_lower = mean(ci_lower),
             ci_upper = mean(ci_upper)) %>% 
   mutate(model = "lstm")
```

Plot comparing persistence to lstm and IDS to ADS

```{r}
ms <- bind_rows(ms_persist, ms_lstm)

f0_pred_plot <- ms %>% 
  ggplot(aes(x = fct_rev(speech_register), y = m, color = fct_rev(model), 
             group = model)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(width = 0.1), 
                   size = 0.8) + 
  geom_line(position = position_dodge(width = 0.1)) + 
  scale_x_discrete(expand = expand_scale(add = 2)) + 
  scale_color_ptol() +
  lims(y = c(0, max(ms$ci_upper) + 0.05)) + 
  labs(x = NULL, y = "Mean Posterior Prob.", color = "Model:") +
  theme(legend.position = "top")

fname <- paste0("writing/figs/", "overall_pred_plot-", config_obj$exp_config$dataset_name, ".png")

save_plot(here(fname), f0_pred_plot, ncol = 0.7, nrow = 1.5)

f0_pred_plot
```

Plot across runs of experiment 

```{r}
d_results %>% 
  ggplot(aes(x = exp_run_id, y = m, color = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(width = 0.1),
                  size = 0.8) +
  ylim(0, NA) +
  labs(x = "Experiment Run ID", y = "Mean Prob.", color = "Speech Register") +
  scale_color_ptol() 
```

### Get model vals to plot

Aggregate at the segement level

```{r}
ss_segment <- d_analysis %>% 
  filter(exp_run_id == "1") %>% 
  group_by(seg_id, speaker_id, speech_register) %>% 
  summarise(m_seg = mean(prob_mass), m_acc_seg = mean(correct_pred))

ss <- ss_segment %>% 
  group_by(speaker_id, speech_register) %>% 
  summarise(m_prob = mean(m_seg), m_acc = mean(m_acc_seg))
```

Now plot

```{r}
delta_val <- 0.9

m_prob_priors <- get_prior(m_seg ~ speech_register + (1 + speech_register | speaker_id), 
                         data = ss_segment, 
                         family = "beta")

m_prob_fit <- brm(m_seg ~ speech_register + (1 + speech_register | speaker_id), 
                 data = ss_segment, 
                 family = "beta",
                 prior = m_prob_priors,
                 control = list(adapt_delta = delta_val), 
                 cores = 4)

post <- posterior_samples(m_prob_fit)
  
m_summary <- post %>% 
    mutate(p_IDS = inv_logit_scaled(b_Intercept),
           p_ADS = inv_logit_scaled(b_Intercept + b_speech_registerADS),
           diff_correct = p_ADS - p_IDS,
           sample_id = 1:nrow(.)) %>% 
    select(sample_id, p_ADS, p_IDS, diff_correct) %>% 
    gather(key = "type", value = "value", -sample_id) %>% 
    group_by(type) %>% 
    median_hdi(.width = 0.9)
```

## Predictability as a function of utterance length

Here, we replicate Rasanen et al. (2018)'s analysis of how the predictability of the intonation changes as a function of position in the utterances, and whether the differences in IDS and ADS would be specific to certain position(s).

Plot the F0 predictability measures as a function of duration of the utterance for both speech registers.

```{r}
d_analysis %>% 
  filter(duration_sec <= 15) %>% 
  group_by(seg_id, speech_register, speaker_id, duration_sec, exp_run_id) %>% 
  summarise(m_prob = mean(prob_mass)) %>% 
  ggplot(aes(x = duration_sec, y = m_prob, color = speech_register)) +
  geom_point(alpha = 0.4, size = 2) +
  geom_smooth(method = "lm", size = 1.5) +
  lims(y = c(0,1)) +
  scale_color_ptol() +
  labs(x = "Utterance Duration (sec)", y = "Mean Prob.",color = "Speech Register") +
  facet_wrap(~ exp_run_id, nrow = 1) +
  theme(legend.position = "top")
```

Collapse across runs

```{r}
duration_plot <- d_analysis %>% 
  filter(duration_sec <= 20) %>% 
  group_by(seg_id, speech_register, speaker_id, duration_sec) %>% 
  summarise(m_prob = mean(prob_mass)) %>% 
  ggplot(aes(x = duration_sec, y = m_prob, color = speech_register)) +
  geom_point(alpha = 0.25, size = 2) +
  geom_smooth(method = "loess", size = 1.5, span = 1) +
  lims(y = c(0,1)) +
  scale_color_ptol() +
  labs(x = "Utterance Duration (sec)", y = "Mean Posterior Prob.",color = "Speech Register:") +
  theme(legend.position = "top")

fname <- paste0("writing/figs/", "duration_plot-", config_obj$exp_config$dataset_name, ".png")
save_plot(here(fname), plot = duration_plot, ncol = 1.5, nrow = 2)

duration_plot
```

## Predictability as a function of within-utterance position

```{r}
too_short_list <- d_analysis %>% 
  ungroup() %>% 
  distinct(seg_id, time_bin_id) %>% 
  count(seg_id) %>% 
  filter(n <= 1) %>% 
  pull(seg_id)

n_bins <- 10

d_within_utt <- d_analysis %>% 
  filter(!(seg_id %in% too_short_list)) %>% 
  group_by(seg_id) %>% 
  mutate(time_bin_within_utt = cut_interval(time_bin_id_num, n_bins, labels = F))

ss_within_utt <- d_within_utt %>% 
  group_by(seg_id, speaker_id, time_bin_within_utt, speech_register, exp_run_id) %>% 
  summarise(m_seg_prob = mean(prob_mass), m_acc_seg = mean(correct_pred)) %>% 
  group_by(speaker_id, time_bin_within_utt, speech_register, exp_run_id) %>% 
  summarise(m_prob = mean(m_seg_prob), m_acc = mean(m_acc_seg)) 

ms_within_utt <- ss_within_utt %>% 
  group_by(time_bin_within_utt, speech_register, exp_run_id) %>% 
  tidyboot_mean(column = m_prob) 

ms_within_utt %>% 
  ggplot(aes(x = as_factor(time_bin_within_utt), y = empirical_stat, 
             color = speech_register, group = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_jitter(0.03)) +
  geom_line(size = 1) +
  scale_color_ptol() +
  lims(y = c(0, NA)) +
  labs(x = "Relative position in Utterance", y = "Mean Prob.", color = "Speech Register") +
  theme(legend.position = "top") +
  facet_wrap(~ exp_run_id, nrow = 1)
```

Collapse across runs

```{r}
ss_within_utt <- d_within_utt %>% 
  group_by(seg_id, speaker_id, time_bin_within_utt, speech_register, exp_run_id) %>% 
  summarise(m_seg_prob = mean(prob_mass),
            m_acc_seg = mean(correct_pred)) %>% 
  group_by(speaker_id, time_bin_within_utt, speech_register) %>% 
  summarise(m_prob = mean(m_seg_prob), 
            m_acc = mean(m_acc_seg)) 

ms_within_utt <- ss_within_utt %>% 
  group_by(time_bin_within_utt, speech_register) %>% 
  tidyboot_mean(column = m_prob) 

temporal_pred_plot <- ms_within_utt %>% 
  ggplot(aes(x = as_factor(time_bin_within_utt), y = empirical_stat, 
             color = speech_register, group = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_jitter(0.03)) +
  geom_label_repel(data = subset(ms_within_utt, 
                                 time_bin_within_utt == max(time_bin_within_utt)),
                   nudge_x = 1,
                   aes(label = speech_register)) +
  geom_line(size = 1) +
  scale_color_ptol() +
  guides(color = F) + 
  lims(y = c(0, NA)) +
  coord_cartesian(xlim = c(0, 11)) +
  labs(x = "Relative position in Utterance", y = "Mean Posterior Prob.", 
       color = "Speech Register") +
  theme(legend.position = "top") 

temporal_pred_plot
```

Split this plot by long vs. short utterances:

```{r}
if(config_obj$exp_config$dataset_name == "ManyBabies") {
  split_val <- 3500
} else {
  split_val <- 2500
}

d_within_utt <- d_within_utt %>% 
  mutate(utt_length_type = ifelse(duration_ms <= split_val, "short", "long"))

ss_within_utt <- d_within_utt %>% 
  group_by(seg_id, speaker_id, time_bin_within_utt, 
           speech_register, utt_length_type, exp_run_id) %>% 
  summarise(m_prob_seg = mean(prob_mass)) %>% 
  group_by(speaker_id, time_bin_within_utt, speech_register, utt_length_type) %>% 
  summarise(m_prob = mean(m_prob_seg)) 
```

```{r}
ms_within_utt <- ss_within_utt %>% 
  group_by(time_bin_within_utt, speech_register, utt_length_type) %>% 
  tidyboot_mean(column = m_prob) 

temporal_plot_split <- ms_within_utt %>% 
  ggplot(aes(x = as_factor(time_bin_within_utt), y = empirical_stat, 
             color = speech_register, group = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(0.1)) +
  geom_label_repel(data = subset(ms_within_utt, time_bin_within_utt == max(time_bin_within_utt)),
                   nudge_x = 1,
                   aes(label = speech_register)) +
  geom_line(size = 1,
            position = position_dodge(0.1)) +
  scale_color_ptol() +
  guides(color = "none") + 
  lims(y = c(0, NA)) +
  coord_cartesian(xlim = c(0, 11)) +
  labs(x = "Relative position in Utterance", y = "Mean Posterior Prob.") +
  theme(legend.position = "right") +
  facet_wrap(~ fct_rev(utt_length_type), ncol = 1)

temporal_plot_split
```

Same plot but collapse across relative position in utterance

```{r}
ms <- d_within_utt %>% 
  group_by(speaker_id, seg_id, speech_register, utt_length_type) %>% 
  summarise(m_seg = mean(prob_mass)) %>% 
  group_by(speaker_id, speech_register, utt_length_type) %>% 
  summarise(m_speaker = mean(m_seg)) %>% 
  group_by(speech_register, utt_length_type) %>% 
  tidyboot_mean(column = m_speaker) 

temporal_plot_collapsed <- ms %>%   
  ggplot(aes(x = fct_rev(utt_length_type), y = empirical_stat, color = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(0.2), size = 0.7) +
  scale_x_discrete(expand = expand_scale(add = 1.5)) +
  ylim(0, NA) +
  labs(x = "Utterance Length", y = "Mean Posterior Prob.", 
       color = "Speech Register:") +
  scale_color_ptol() +
  theme(legend.position = "top")

temporal_plot_collapsed
```

Pull plots together and save

```{r}
temporal_final_plot <- plot_grid(temporal_pred_plot, temporal_plot_collapsed, 
          scale = c(0.95, 0.85), 
          rel_widths = c(1, 0.8))

fname <- paste0("writing/figs/", "temporal_pred_plot-", config_obj$exp_config$dataset_name, ".png")
save_plot(here(fname), plot = temporal_final_plot, ncol = 2, nrow = 1.7)

temporal_final_plot
```

## Predictability as a function of CDS in training

```{r}
ss_prop_train <- d %>% 
  filter(is_target_cluster == TRUE) %>% 
  group_by(seg_id, speech_register, prop_cds_train, speaker_id, duration_ms) %>% 
  summarise(m_prob = mean(prob_mass),
            m_acc = mean(correct_pred))

ms_prop_train <- ss_prop_train %>% 
  group_by(prop_cds_train, speech_register) %>% 
  tidyboot_mean(column = m_prob) 

ms_prop_train %>% 
  ggplot(aes(x = as_factor(prop_cds_train), y = empirical_stat, 
             fill = speech_register, group = speech_register)) +
  geom_col(position = 'dodge', width = 0.3) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size =1,
                 position=position_dodge(width=0.3)) +
  scale_fill_ptol() +
  lims(y = c(0, NA)) +
  labs(x = "Prop. CDS in Training Data", y = "Mean Prob. Correct") +
  theme(legend.position = "top")
```

## Error analysis

Plot the distribution of success and errors with the cluster shapes

```{r, fig.width=8, out.width="95%"}
d_kmeans <- read_rds(here(paste0(config_obj$paths_config$pitch_sum_path, 
                                 "lena-pred-clustering-outputs-", 
                                 config_obj$exp_config$dataset_name, 
                                  "-",
                                 config_obj$kmeans_config$scale_coefs, 
                                 ".rds")))
shape_val <- "shapes_12-run2"
d_shapes <- d_kmeans[[shape_val]]
cluster_shapes <- plot_cluster_shapes(d_shapes$centers, scaled = TRUE)

target_dist_plot <-  d %>% 
  filter(is_target_cluster == T) %>% 
  ggplot(aes(as_factor(target_cluster), fill = speech_register)) +
  geom_histogram(stat='count') +
  labs(x = "Cluster Category", title = "Target Distribution") +
  facet_wrap(~speech_register, ncol = 1) +
  scale_fill_ptol() +
  guides(fill = F) + 
  theme(legend.position = "bottom")

errors_plot <- d %>% 
  filter(is_target_cluster == T) %>% 
  ggplot(aes(as_factor(predicted_cluster), fill = speech_register)) +
  geom_histogram(stat='count') +
  labs(x = "Cluster Category", title = "Model predictions") +
  facet_wrap(speech_register~correct_pred, ncol = 2) +
  scale_fill_ptol() +
  guides(fill = F) + 
  theme(legend.position = "bottom")

plot_grid(cluster_shapes, target_dist_plot, errors_plot,
          nrow = 1,
          rel_widths =  c(1.1, 1, 1.2),
          scale = c(1, 0.8, 0.8))
```

## IDSLabel-specific Analyses

Join participant metadata

```{r}
idslab_meta <- read_csv(here("data/00_metadata/IDS_label/", "ACLEW_list_of_corpora-restricted20171031.csv")) %>% 
  rename(speaker_id = ID) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)


vd_demo <- read_csv(here("data/00_metadata/IDS_label/", 
                         "vd_demo_update-restricted20171031.csv")) %>% 
  rename(speaker_id = ID) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)

chi_ages <- read_csv(here("data/00_metadata/IDS_label/", 
                          "proposed_sample_summary-22Jun16 with substitutes-FINALbeforeIDSlabelannotation-restricted20171031.csv")) %>% 
  rename(speaker_id = `Sample ID`,
         chi_gender = Gender) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)

full_meta_idslab <- bind_rows(idslab_meta, vd_demo, chi_ages) %>% distinct()

process_mcdiv_id <- function(s) {
  if (str_length(s) > 6) {
    paste(str_split(s, "-", simplify = T)[1],
      str_split(s, "-", simplify = T)[2],
      sep = "-")
  } else {
    s
  }
}

ss$speaker_id <- map_chr(ss$speaker_id, process_mcdiv_id)

ss <- ss %>% left_join(full_meta_idslab, by = "speaker_id")
```

Individual differences in predictability of IDS vs. ADS

```{r}
get_register_diff <- function(x) {
  ads_val <- x[2]
  ids_val <- x[1]
  ads_val - ids_val
}

ss <- ss %>% 
  group_by(speaker_id) %>% 
  mutate(reg_diff = get_register_diff(m_prob))

ss %>% 
  filter(!is.na(reg_diff)) %>% 
  distinct(speaker_id, reg_diff) %>% 
  ggplot(aes(x = reg_diff, y = fct_reorder(speaker_id, reg_diff))) +
  geom_point(size = 4, alpha = 0.5) +
  lims(x = c(-0.3, 0.3)) +
  theme(axis.text.x = element_text(angle = 90)) +
  geom_vline(xintercept = 0, lty = "dashed") +
  labs(x = "Predictability Difference \n (Larger value means ADS more predictable)", y = "Child ID")
```

Predictability of both registers a function of age

```{r}
ss %>% 
  ggplot(aes(x = AgeMonths, y = m_prob, color = speech_register)) +
  geom_point(size = 2.5, alpha = 0.5) +
  labs(x = "Child's Age (months)", y = "Mean Prob.", color = "Speech Register") + 
  geom_smooth(method = "lm", se = F, size = 2) +
  scale_color_ptol() +
  theme(legend.position = "top")
```

Diff in predictability as a function of age

```{r}
ss %>% 
  ggplot(aes(x = AgeMonths, y = reg_diff)) +
  geom_point(size = 2.5, alpha = 0.5) +
  labs(x = "Child's Age (months)", y = "Predictability Difference (ADS vs. IDS)") + 
  geom_smooth(method = "lm", se = F, size = 2) +
  scale_color_ptol() +
  theme(legend.position = "top")
```

## Sample some high vs. low predictability utterances

```{r}
ss_seg <- d_analysis %>% 
  filter(exp_run_id == "1") %>% 
  group_by(seg_id, speaker_id, speech_register, duration_sec) %>% 
  summarise(m_seg = mean(prob_mass), m_acc_seg = mean(correct_pred)) %>% 
  distinct(seg_id, m_seg, duration_sec) 
  
ss_seg$pred_quartile = ntile(ss_seg$m_seg, n = 4)  
```

```{r}
samples <- ss_seg %>% 
  ungroup() %>% 
  group_by(pred_quartile, speech_register) %>% 
  sample_n(1)

samples
```
