---
title: "Lena-pred Data Visualization"
output: html_document
---

## Set up

```{r libraries and config, warning=FALSE, include=FALSE}
source(here::here("code/00_config/lena-pred-libraries.R"))
source(here("code/00_config/lena-pred-config.R"))
# TODO: fix extract_preds() function to handle new results format with folds
d <- read_lena_pred_data("lena-pred-lstm-preds", config_obj, "rds", is_pitch = FALSE)
d_persist <- read_lena_pred_data("lena-pred-persist-results", config_obj, "csv")

if (config_obj$exp_config$dataset_name == "IDSLabel") {
  d <- d %>% 
    mutate(speaker_id = str_split(seg_id, "_", simplify = TRUE)[1])  
}
```

## Baseline persistence model results. 

```{r}
d_persist %>% 
  ggplot(aes(x = as_factor(n_qshapes), y = acc, 
             color = as_factor(prop_cds),
             group = prop_cds)) +
  geom_point(size = 2) +
  geom_line(size = 1) +
  scale_color_ptol() +
  labs(x = "Num of Q-shapes", y = "Mean Accuracy", 
       color = "Prop. CDS in Dataset",
       title = "Persistence Model")
```

## ADS vs. CDS average predictability

Replicate the key analysis in Rasanen et al. (2018), which compared the average predictability of ADS vs. CDS.

First we aggregate the results for each run of the experiment for only the 50/50 split of CDS in the training data.

```{r}
d_analysis <- d %>% 
  filter(prop_cds_train == 0.5, is_target_cluster == TRUE) %>% 
  mutate(exp_run_id = str_remove(exp_run_id, "run"),
         duration_ms = as.numeric(duration_ms),
         duration_sec = duration_ms / 1000,
         time_bin_id_num = as.numeric(time_bin_id)) 

d_results <- d_analysis %>% 
  group_by(exp_run_id) %>% 
  nest() %>% 
  mutate(results = map(data, analyze_one_exp, metric = m_acc))
```

Now make the key plots:

```{r}
d_results %>% 
  select(exp_run_id, results) %>% 
  unnest() %>% 
  ggplot(aes(x = exp_run_id, y = empirical_stat, color = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(width = 0.1),
                  size = 0.8) +
  ylim(0, 0.6) +
  labs(x = "Experiment Run ID", y = "Mean Prob.", color = "Speech Register") +
  scale_color_ptol() 
```

```{r}
ss_segment <- d_analysis %>% 
  filter(exp_run_id == "1") %>% 
  group_by(seg_id, speaker_id, speech_register) %>% 
  summarise(m_seg = mean(prob_mass), m_acc_seg = mean(correct_pred))

ss <- ss_segment %>% 
  group_by(speaker_id, speech_register) %>% 
  summarise(m_prob = median(m_seg), m_acc = mean(m_acc_seg))
```

```{r}
delta_val <- 0.9

m_prob_priors <- get_prior(m_seg ~ speech_register + (1 + speech_register | speaker_id), 
                         data = ss_segment, 
                         family = "beta")

m_prob_fit <- brm(m_seg ~ speech_register + (1 + speech_register | speaker_id), 
                 data = ss_segment, 
                 family = "beta",
                 prior = m_prob_priors,
                 control = list(adapt_delta = delta_val), 
                 cores = 4)

post <- posterior_samples(m_prob_fit)
  
m_summary <- post %>% 
    mutate(p_IDS = inv_logit_scaled(b_Intercept),
           p_ADS = inv_logit_scaled(b_Intercept + b_speech_registerADS),
           diff_correct = p_ADS - p_IDS,
           sample_id = 1:nrow(.)) %>% 
    select(sample_id, p_ADS, p_IDS, diff_correct) %>% 
    gather(key = "type", value = "value", -sample_id) %>% 
    group_by(type) %>% 
    median_hdi(.width = 0.9)
```

```{r eval=FALSE, include=FALSE}
plot_model_output(mb_ms, col_name = m_prob, y_axis_lab = "Mean Prob. Mass")
```

## Predictability as a function of utterance length

Here, we replicate Rasanen et al. (2018)'s analysis of how the predictability of the intonation changes as a function of position in the utterances, and whether the differences in IDS and ADS would be specific to certain position(s).

Plot the F0 predictability measures as a function of duration of the utterance for both speech registers.

```{r}
d_analysis %>% 
  filter(duration_sec <= 15) %>% 
  group_by(seg_id, speech_register, speaker_id, duration_sec, exp_run_id) %>% 
  summarise(m_prob = mean(prob_mass)) %>% 
  ggplot(aes(x = duration_sec, y = m_prob, color = speech_register)) +
  geom_point(alpha = 0.4, size = 2) +
  geom_smooth(method = "lm", size = 1.5) +
  lims(y = c(0,1)) +
  scale_color_ptol() +
  labs(x = "Utterance Duration (sec)", y = "Mean Prob.",color = "Speech Register") +
  facet_wrap(~ exp_run_id, nrow = 1) +
  theme(legend.position = "top")
```

## Predictability as a function of within-utterance position

```{r}
too_short_list <- d_analysis %>% 
  ungroup() %>% 
  distinct(seg_id, time_bin_id) %>% 
  count(seg_id) %>% 
  filter(n <= 1) %>% 
  pull(seg_id)

n_bins <- 10

d_within_utt <- d_analysis %>% 
  filter(!(seg_id %in% too_short_list)) %>% 
  group_by(seg_id, exp_run_id) %>% 
  mutate(time_bin_within_utt = cut_interval(time_bin_id_num, n_bins, labels = F))

ss_within_utt <- d_within_utt %>% 
  group_by(seg_id, speaker_id, time_bin_within_utt, speech_register, exp_run_id) %>% 
  summarise(m_seg_prob = mean(prob_mass), m_acc_seg = mean(correct_pred)) %>% 
  group_by(speaker_id, time_bin_within_utt, speech_register, exp_run_id) %>% 
  summarise(m_prob = mean(m_seg_prob), m_acc = mean(m_acc_seg)) 

ms_within_utt <- ss_within_utt %>% 
  group_by(time_bin_within_utt, speech_register, exp_run_id) %>% 
  tidyboot_mean(column = m_prob) 

ms_within_utt %>% 
  ggplot(aes(x = as_factor(time_bin_within_utt), y = empirical_stat, 
             color = speech_register, group = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_jitter(0.03)) +
  geom_line(size = 1) +
  scale_color_ptol() +
  lims(y = c(0, NA)) +
  labs(x = "Relative position in Utterance", y = "Mean Prob.", color = "Speech Register") +
  theme(legend.position = "top") +
  facet_wrap(~ exp_run_id, nrow = 1)
```

Split this plot by long vs. short utterances:

```{r}
split_val <- 1000

d_within_utt <- d_within_utt %>% 
  mutate(utt_length_type = ifelse(duration_ms <= split_val, "short", "long"))

ss_within_utt <- d_within_utt %>% 
  group_by(seg_id, speaker_id, time_bin_within_utt, speech_register, 
           utt_length_type, exp_run_id) %>% 
  summarise(m_prob_seg = mean(prob_mass)) %>% 
  group_by(speaker_id, time_bin_within_utt, speech_register, utt_length_type, exp_run_id) %>% 
  summarise(m_prob = mean(m_prob_seg)) 

ms_within_utt <- ss_within_utt %>% 
  group_by(time_bin_within_utt, speech_register, utt_length_type, exp_run_id) %>% 
  tidyboot_mean(column = m_prob) 

ms_within_utt %>% 
  ggplot(aes(x = as_factor(time_bin_within_utt), y = empirical_stat, 
             color = speech_register, group = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper), 
                  position = position_dodge(0.1)) +
  geom_line(size = 1,
            position = position_dodge(0.1)) +
  scale_color_ptol() +
  guides(color = "none") + 
  lims(y = c(0, NA)) +
  labs(x = "Relative position in Utterance", y = "Mean Prob.", color = "Speech Register") +
  theme(legend.position = "right") +
  facet_grid(fct_rev(utt_length_type) ~ exp_run_id)
```

```{r}
ms <- d_within_utt %>% 
  group_by(speaker_id, seg_id, speech_register, utt_length_type, exp_run_id) %>% 
  summarise(m_seg = mean(prob_mass)) %>% 
  group_by(speaker_id, speech_register, utt_length_type, exp_run_id) %>% 
  summarise(m_speaker = mean(m_seg)) %>% 
  group_by(speech_register, utt_length_type, exp_run_id) %>% 
  tidyboot_mean(column = m_speaker) 

ms %>%   
  ggplot(aes(x = fct_rev(utt_length_type), y = empirical_stat, color = speech_register)) +
  geom_pointrange(aes(ymin = ci_lower, ymax = ci_upper),
                  position = position_dodge(0.1),
                  size = 0.8) +
  ylim(0, 0.4) +
  guides(color = "none") +
  labs(x = "Utterance Length", y = "Mean Prob.", color = "Speech Register") +
  scale_color_ptol() +
  facet_wrap(~exp_run_id)
```

## Predictability as a function of CDS in training

```{r}
ss_prop_train <- d %>% 
  filter(is_target_cluster == TRUE) %>% 
  group_by(seg_id, speech_register, prop_cds_train, speaker_id, duration_ms) %>% 
  summarise(m_prob = mean(prob_mass),
            m_acc = mean(correct_pred))

ms_prop_train <- ss_prop_train %>% 
  group_by(prop_cds_train, speech_register) %>% 
  tidyboot_mean(column = m_prob) 

ms_prop_train %>% 
  ggplot(aes(x = as_factor(prop_cds_train), y = empirical_stat, 
             fill = speech_register, group = speech_register)) +
  geom_col(position = 'dodge', width = 0.3) + 
  geom_linerange(aes(ymin = ci_lower, ymax = ci_upper), size =1,
                 position=position_dodge(width=0.3)) +
  scale_fill_ptol() +
  lims(y = c(0, NA)) +
  labs(x = "Prop. CDS in Training Data", y = "Mean Prob. Correct") +
  theme(legend.position = "top")
```

## Error analysis

Plot the distribution of success and errors with the cluster shapes

```{r, fig.width=8, out.width="95%"}
d_kmeans <- read_rds(here(paste0(config_obj$paths_config$pitch_sum_path, 
                                 "lena-pred-kmeans-outputs-", 
                                 config_obj$exp_config$dataset_name, ".rds")))
shape_val <- "shapes_12-run1"
d_shapes <- d_kmeans[[shape_val]]
cluster_shapes <- plot_cluster_shapes(d_shapes$centers, scaled = TRUE)

target_dist_plot <-  d %>% 
  filter(is_target_cluster == T) %>% 
  ggplot(aes(as_factor(target_cluster), fill = speech_register)) +
  geom_histogram(stat='count') +
  labs(x = "Cluster Category", title = "Target Distribution") +
  facet_wrap(~speech_register, ncol = 1) +
  scale_fill_ptol() +
  guides(fill = F) + 
  theme(legend.position = "bottom")

errors_plot <- d %>% 
  filter(is_target_cluster == T) %>% 
  ggplot(aes(as_factor(predicted_cluster), fill = speech_register)) +
  geom_histogram(stat='count') +
  labs(x = "Cluster Category", title = "Model predictions") +
  facet_wrap(speech_register~correct_pred, ncol = 2) +
  scale_fill_ptol() +
  guides(fill = F) + 
  theme(legend.position = "bottom")

plot_grid(cluster_shapes, target_dist_plot, errors_plot,
          nrow = 1,
          rel_widths =  c(1.1, 1, 1.2),
          scale = c(1, 0.8, 0.8))
```


## IDSLabel-specific Analyses

Join participant metadata

```{r}
idslab_meta <- read_csv(here("data/00_metadata/IDS_label/", "ACLEW_list_of_corpora-restricted20171031.csv")) %>% 
  rename(speaker_id = ID) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)


vd_demo <- read_csv(here("data/00_metadata/IDS_label/", "vd_demo_update-restricted20171031.csv")) %>% 
  rename(speaker_id = ID) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)

chi_ages <- read_csv(here("data/00_metadata/IDS_label/", 
                          "proposed_sample_summary-22Jun16 with substitutes-FINALbeforeIDSlabelannotation-restricted20171031.csv")) %>% 
  rename(speaker_id = `Sample ID`,
         chi_gender = Gender) %>% 
  select(speaker_id, Corpus, AgeMonths, chi_gender)

full_meta_idslab <- bind_rows(idslab_meta, vd_demo, chi_ages) %>% distinct()

process_mcdiv_id <- function(s) {
  if (str_length(s) > 6) {
    paste(str_split(s, "-", simplify = T)[1],
      str_split(s, "-", simplify = T)[2],
      sep = "-")
  } else {
    s
  }
}

ss$speaker_id <- map_chr(ss$speaker_id, process_mcdiv_id)

ss <- ss %>% left_join(full_meta_idslab, by = "speaker_id")
```

Individual differences in predictability of IDS vs. ADS

```{r}
get_register_diff <- function(x) {
  ads_val <- x[2]
  ids_val <- x[1]
  ads_val - ids_val
}

ss <- ss %>% 
  group_by(speaker_id) %>% 
  mutate(reg_diff = get_register_diff(m_prob))

ss %>% 
  filter(!is.na(reg_diff)) %>% 
  distinct(speaker_id, reg_diff) %>% 
  ggplot(aes(x = reg_diff, y = fct_reorder(speaker_id, reg_diff))) +
  geom_point(size = 4, alpha = 0.5) +
  lims(x = c(-0.3, 0.3)) +
  theme(axis.text.x = element_text(angle = 90)) +
  coord_flip() +
  geom_vline(xintercept = 0, lty = "dashed") +
  labs(x = "Predictability Difference \n (Larger value means ADS more predictable)", y = "Child ID")
```

Predictability of both registers a function of age

```{r}
ss %>% 
  ggplot(aes(x = AgeMonths, y = m_prob, color = speech_register)) +
  geom_point(size = 2.5, alpha = 0.5) +
  labs(x = "Child's Age (months)", y = "Mean Prob.", color = "Speech Register") + 
  geom_smooth(method = "lm", se = F, size = 2) +
  scale_color_ptol() +
  theme(legend.position = "top")
```

Diff in predictability as a function of age

```{r}
ss %>% 
  ggplot(aes(x = AgeMonths, y = reg_diff)) +
  geom_point(size = 2.5, alpha = 0.5) +
  labs(x = "Child's Age (months)", y = "Predictability Difference (ADS vs. IDS)") + 
  geom_smooth(method = "lm", se = F, size = 2) +
  scale_color_ptol() +
  theme(legend.position = "top")
```

Predictability over the course of the day

Predictability over the course of a conversational block

## Sample some high vs. low predictability utterances

```{r}
ss_seg <- d_analysis %>% 
  filter(exp_run_id == "1") %>% 
  group_by(seg_id, speaker_id, speech_register, duration_sec) %>% 
  summarise(m_seg = mean(prob_mass), m_acc_seg = mean(correct_pred)) %>% 
  distinct(seg_id, m_seg, duration_sec) 
  
ss_seg$pred_quartile = ntile(ss_seg$m_seg, n = 4)  
```

```{r}
samples <- ss_seg %>% 
  ungroup() %>% 
  group_by(pred_quartile, speech_register) %>% 
  sample_n(1)

samples
```



