---
title: "Lena-pred Data Analysis Models"
output: html_document
---

## Modeling set up

To compare the output of the LSTM between CDS and ADS, we will use the brms (Burkner, 2017) R package to fit Bayesian mixed-effects regression models. The mixed-effects approach will allow us to model the nested structure of our data – multiple recordings for each child and for each type of utterance (CDS vs. ADS). We will use Bayesian estimation to quantify uncertainty in our estimates, which we communicate using a 95% Highest Density Interval (HDI), providing a range of credible values for the differences between CDS and ADS, given the data and model. 

```{r libraries and config}
source(here::here("code/00_config/lena-pred-libraries.R"))
source(here::here("code/00_config/lena-pred-pitch-config.R"))
source(here("code/00_helper_functions/lstm-helpers.R"))
source(here("code/00_helper_functions/plot_helpers.R"))
```

Read model predictions 

```{r}
data_path <- paste0("data/03_summaries/", dataset_config$data_set)
d_list <- read_rds(here(data_path, "lena-pred-lstm-preds.rds"))
```

```{r}
delta_val <- 0.9
```

```{r}
get_speaker_id_mb <- function(s) {str_extract(s, "(?<=_?)\\d+")}

d <- extract_preds(d_list)

d <- d %>% mutate(speaker_id = get_speaker_id_mb(seg_id),
                  duration_ms = as.numeric(duration_ms)) 
```

Aggregate accuracy and posterior probability over correct cluster:

```{r}
seg_ms <- d %>% 
  filter(is_target_cluster == TRUE) %>% 
  group_by(seg_id, speech_register, n_qshapes, speaker_id, duration_ms) %>% 
  summarise(m_prob = mean(prob_mass),
            m_acc = mean(correct_pred)) %>% 
  group_by(seg_id, speech_register, speaker_id, duration_ms) %>% 
  summarise(m_prob = mean(m_prob),
            m_acc = mean(m_acc)) 

# add scaled duration value
seg_ms$duration_scaled <- scale(seg_ms$duration_ms)
```

## ManyBabies

### ADS vs. CDS average predictability

Output variable is the probability of correct F0 prediction, which is bound between [0, 1], so we use Beta regression. The key prediction is that th probability of correct F0 prediction will be lower for CDS compared to ADS, indicating that the LSTM put a lower posterior probability on the correct F0 trajectory when processing CDS.

```{r beta reg}
m_prob_priors <- get_prior(m_prob ~ speech_register + (1 + speech_register|speaker_id), 
                         data = seg_ms, 
                         family = "beta")

m_prob_fit <- brm(m_prob ~ speech_register + (1 + speech_register|speaker_id), 
                 data = seg_ms, 
                 family = "beta",
                 prior = m_prob_priors,
                 control = list(adapt_delta = delta_val))

results_obj <- list(prob_model = get_bda_results(m_prob_fit))
```

Also analyze perplexity.

```{r gamma reg, include=FALSE}
m_perp_priors <- get_prior(m_perp ~ speech_register + (1 + speech_register|speaker) + 
                             (1 + speech_register | utt_type), 
                         data = d, 
                         family = "Gamma")

m_perp_fit <- brm(m_perp ~ speech_register + (1 + speech_register|speaker) + (1 + speech_register | utt_type), 
                 data = d, 
                 family = "Gamma",
                 prior = m_perp_priors)
```

### Accuracy predicting next cluster

Classification accuracy, which is a binary outcome, so we use logistic regression, The key prediction is that accuracy will be lower for CDS compared to ADS.

```{r logistic reg}
d_logit <- d %>% 
  filter(n_qshapes == 16) %>% # select just one dataset of q-shapes
  distinct(seg_id, speaker_id, speech_register, time_bin_id, 
           correct_pred, target_cluster) %>% 
  group_by(speaker_id, speech_register) %>% 
  summarise(n_correct = sum(correct_pred),
            n_bins = n(),
            prop_correct = n_correct / n_bins)


m_acc_priors <- get_prior(n_correct | trials(n_bins) ~ speech_register + (1 + speech_register|speaker_id),
                         data = d_logit, 
                         family = "binomial")

m_acc_fit <- brm(n_correct | trials(n_bins) ~ speech_register + (1 + speech_register|speaker_id),
                 data = d_logit, 
                 family = "binomial",
                 prior = m_acc_priors,
                 control = list(adapt_delta = delta_val))

results_obj$acc_model <- get_bda_results(m_acc_fit)
```

### Predictability as a function of utterance length

Within- utterance temporal predictability. The probability of correct F0 prediction as a function of utterance duration. A probability, bound between [0, 1], so we again use Beta regression. We predict an interaction: that the probability of correct F0 prediction for ADS will increase *more* as a function of utterance duration as compared to CDS, indicating that CDS is harder to predict later in an utterance.

```{r beta reg duration}
m_prob_priors <- get_prior(m_prob ~ speech_register * duration_scaled + (1 + speech_register|speaker_id), 
                         data = seg_ms, 
                         family = "beta")

m_prob_fit <- brm(m_prob ~ speech_register * duration_scaled + (1 + speech_register|speaker_id), 
                  data = seg_ms, 
                  family = "beta",
                  prior = m_prob_priors,
                  control = list(adapt_delta = delta_val))

```

```{r}

get_bda_results(m_prob_fit, d = seg_ms, mod_type = "duration") -> blah


get_bda_results <- function(m_fit, d, mod_type) {
  
  post <- posterior_samples(m_fit)
  
  if (mod_type=="duration") {
    dur_m <- mean(d$duration_ms)
    dur_sd <- sd(d$duration_ms)
    
    m_summary <- post %>% 
      mutate(p_ADS = inv_logit_scaled(b_Intercept),
             p_IDS = inv_logit_scaled(b_Intercept + b_speech_registerIDS),
             diff_ADS_IDS = p_ADS - p_IDS,
             b_duration_ms = b_duration_scaled * dur_sd + dur_m,
             b_dur_int_ms = b_speech_registerIDS:duration_scaled * dur_sd + dur_m,
             sample_id = 1:nrow(.)) %>%  
      select(sample_id, p_ADS, p_IDS, diff_ADS_IDS, b_duration_ms, b_dur_int_ms) %>% 
      gather(key = "type", value = "value", -sample_id) %>% 
      group_by(type) %>% 
      median_hdi(.width = 0.95)
    
  } else {
    m_summary <- post %>% 
      mutate(p_ADS = inv_logit_scaled(b_Intercept),
             p_IDS = inv_logit_scaled(b_Intercept + b_speech_registerIDS),
             diff_ADS_IDS = p_ADS - p_IDS,
             sample_id = 1:nrow(.)) 
    select(sample_id, p_ADS, p_IDS, diff_ADS_IDS) %>% 
      gather(key = "type", value = "value", -sample_id) %>% 
      group_by(type) %>% 
      median_hdi(.width = 0.95)
  }
  
  model_results <- list(post_samples = post, summary = m_summary)
}

```

Save model outputs

```{r}
write_rds(results_obj, here("lena-pred-bda-results.rds"))
```


## IDSLabel

### Set up

### ADS vs. CDS average predictability

### Accuracy predicting next cluster

### Predictability as a function of utterance length


